---
title: "Baseball5_TY"
author: "Thomas Young"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(randomForest)
library(regclass) 
library(rpart) 
library(rpart.plot) 
```

## Load Hitters DATA

Hitters Data Set from Kaggle: <https://www.kaggle.com/datasets/floser/hitters>

```{r cars}
hitters <- read_csv("Hitters.csv")

# clean
hitters <- hitters |>  drop_na()

# Create variable for log salary 
hitters <- hitters |>  mutate(LogSalary = log(Salary))
```

## Build models 

```{r}
set.seed(123)
tree_model <- rpart(LogSalary ~ ., data = hitters, 
                   control = rpart.control(cp = 0.01))

rf_model <- randomForest(LogSalary ~ ., data = hitters)

# Visualize the full tree
rpart.plot(tree_model, extra = 101) # extra=101 for node labels
print(rf_model)
```

## Check Performance

```{r}
# CART R-squared
r_squared <- 1 - (sum((hitters$LogSalary - predict(tree_model, hitters))^2) / 
                   sum((hitters$LogSalary - mean(hitters$LogSalary))^2))

# Random forest R-squared
r_squared_rf <- 1 - (sum((hitters$LogSalary - predict(rf_model, hitters))^2) / 
                       sum((hitters$LogSalary - mean(hitters$LogSalary))^2))

# Compare R-squared values
print(paste("Regression Tree R-squared:", r_squared))
print(paste("Random Forest R-squared:", r_squared_rf))
```

```{r}
# CART RMSE 
rmse <- sqrt(mean((hitters$LogSalary - predict(tree_model, hitters))^2))

# # Random forest RMSE
rmse_rf <- sqrt(mean((hitters$LogSalary - predict(rf_model, hitters))^2))

# Compare RMSE values
print(paste("Regression Tree RMSE:", rmse)) 
print(paste("Random Forest RMSE:", rmse_rf))
```

## Interpretation 

Lower RMSE is better: RMSE (Root Mean Squared Error) measures the average difference between the predicted values and the actual values. The random forest model had a lower RMSE indicates that the model's predictions are closer to the true values.

Improvement in R-squared: The random forest model had a higher R-squared value than the regression tree, suggesting that it explains more of the variance in LogSalary
