---
title: "Baseball4_TY"
author: "Thomas Young"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(regclass) # to visualize the model
library(rpart) # to fit a decision tree
library(rpart.plot) # For visualizing the tree
```

## Load Hitters DATA

Hitters Data Set from Kaggle: <https://www.kaggle.com/datasets/floser/hitters>

```{r cars}
hitters <- read_csv("Hitters.csv")

# clean
hitters <- hitters |>  drop_na()

# Create variable for log salary 
hitters <- hitters |>  mutate(LogSalary = log(Salary))
```

## Build with full data set 

```{r}
set.seed(123)
tree_model <- rpart(LogSalary ~ ., data = hitters, 
                   control = rpart.control(cp = 0.01))

# Visualize the full tree
rpart.plot(tree_model, extra = 101) # extra=101 for node labels
```

```{r}
summary(tree_model)

# Performance
r_squared <- 1 - (sum((hitters$LogSalary - predict(tree_model, hitters))^2) / 
                   sum((hitters$LogSalary - mean(hitters$LogSalary))^2))
print(r_squared)
```
## Interpretation 

- The tree begins with 263 observations (players) and the root node (Node 1) has a predicted average LogSalary of 5.927.
- The primary split is based on Salary < 333.3335.
  - If a player's salary is less than 333.3335, they go to Node 2.
  - If their salary is greater than or equal to 333.3335, they go to Node 3.
  
Branch 1: Lower Salaries (Node 2, Node 4, Node 5, Node 10, Node 11)

Node 2: This node contains 112 players with lower salaries (predicted LogSalary = 5.048).
- Split: The next split is based on Salary < 147.5.
  - If salary is less than 147.5, go to Node 4.
  - If salary is greater than or equal to 147.5, go to Node 5.
  
Node 4: This is a terminal node with 49 players and a predicted LogSalary of 4.581. These are likely players with very low salaries.

Node 5: This node contains 63 players with slightly higher salaries (predicted LogSalary = 5.410).
- Split: Further split based on Salary < 220.
  - If salary is less than 220, go to Node 10.
  - If salary is greater than or equal to 220, go to Node 11.

Node 10: Terminal node with 30 players and predicted LogSalary = 5.202.

Node 11: Terminal node with 33 players and predicted LogSalary = 5.599.

Branch 2: Higher Salaries (Node 3, Node 6, Node 7, Node 12, Node 13, Node 14, Node 15)

Node 3: This node contains 151 players with higher salaries (predicted LogSalary = 6.579).
- Split: Based on Salary < 807.5.
  - If salary is less than 807.5, go to Node 6.
  - If salary is greater than or equal to 807.5, go to Node 7.
  
Node 6: 98 players (predicted LogSalary = 6.313).
- Split: Based on Salary < 542.5.
  - If salary is less than 542.5, go to Node 12.
  - If salary is greater than or equal to 542.5, go to Node 13.
  
Node 12: Terminal node with 46 players and predicted LogSalary = 6.071.

Node 13: Terminal node with 52 players and predicted LogSalary = 6.528.

Node 7: 53 players (predicted LogSalary = 7.072).
- Split: Based on Salary < 1400.
  - If salary is less than 1400, go to Node 14.
  - If salary is greater than or equal to 1400, go to Node 15.
  
Node 14: Terminal node with 40 players and predicted LogSalary = 6.922.

Node 15: Terminal node with 13 players and predicted LogSalary = 7.534. These are likely players with the highest salaries.

The tree primarily uses the Salary variable to make splits, suggesting it's a strong predictor of LogSalary. As you move down the branches, the predicted LogSalary values generally increase, reflecting the salary-based splits

Overall performance of this model was 97% 
