---
title: "Baseball3_TY"
author: "Thomas Young"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(regclass) # to visualize the model
library(rpart) # to fit a decision tree
library(rpart.plot) # For visualizing the tree
```

## Load Hitters DATA

Hitters Data Set from Kaggle: <https://www.kaggle.com/datasets/floser/hitters>

```{r cars}
hitters <- read_csv("Hitters.csv")

# clean
hitters <- hitters |>  drop_na()

# Create variable for log salary 
hitters <- hitters |>  mutate(LogSalary = log(Salary))
```

## Regression Example

```{r}
tree_model <- rpart(LogSalary ~ Years + Hits, data = hitters, 
                   control = rpart.control(maxdepth = 2, cp = 0.01)) 
visualize_model(tree_model)
```

## Compare to another model

```{r}
mod1 <- lm(LogSalary ~ Years + Hits, data = hitters)
summary(mod1)
```

## Compare predictions

```{r}
predictions <- predict(tree_model, newdata = hitters) 

# Add predictions to df:
hitters$PredictedLogSalary <- predictions

# Compute R2
R2 <- 1 - (sum((hitters$LogSalary - predictions)^2) / sum((hitters$LogSalary - mean(hitters$LogSalary))^2))
R2
```

## Interpretation

Root Node (Years < 5): This is the starting point of the tree. It splits the data based on whether the number of years played (Years) is less than 5 or not.

- Yes Branch (Years < 5): If a player has played less than 5 years, the tree follows this branch.
- No Branch (Years >= 5): If a player has played 5 or more years, the tree follows this branch

Left Branch of Root Node (Years < 5):

Node (Years < 4): This node further splits the players who have played less than 5 years based on whether they have played less than 4 years

- Left Leaf (Circle with 4.89, n=62): This represents a terminal node (leaf) where the prediction is made. The value 4.89 is the mean log salary for the 62 players who have played less than 4 years.

- Right Leaf (Circle with 5.58, n=28): This represents another terminal node where the prediction is made. The value 5.58 is the mean log salary for the 28 players who have played between 4 and 5 years (Years >= 4 and Years < 5).

Right Branch of Root Node (Years >= 5):

Node (Hits < 118): This node splits the players who have played 5 or more years based on whether they had less than 118 hits in the previous year.

- Left Leaf (Circle with 8, n=90): This represents a terminal node where the prediction is made. The value 8 is the mean log salary for the 90 players who have played 5 or more years and had less than 118 hits. This could potentially represent outlier data (very high salary for a relatively low number of hits.) It would be worth investigating these data points.

- Right Leaf (Circle with 6.47, n=83): This represents another terminal node where the prediction is made. The value 6.47 is the mean log salary for the 83 players who have played 5 or more years and had 118 or more hits.

Overall Performance of this model was 60%, which was expected as we used only two vairbales.
